Evaluation of clustering parameters - SCPCS000168

### 2024-08-07 analyses/cell-type-wilms-tumor-06/01-clustering_SCPCS000168.Rmd
add 01-clustering abalysis

---
title: "Evaluation of clustering parameters - SCPCS000169"
author: maud-p
date: '`r format(Sys.time(), "%B %d, %Y %H:%M:%S %Z")`'
output:
  html_document: 
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: pygments
    df_print: kable
params:
  sample_name: ~
  sample_path: ~
  out_rds_path: ~
---

@@

In this notebook, we will set up parameters in the seurat workflow 
[normalization --> reduction --> clustering] 
for one Wilms tumor sample (SCPCS000169) of the Wilms Tumor dataset (SCPCP000006).

@@

library(Azimuth)     # For label transfer
```

@@

```{r cfg}
cfg = list()
cfg$padj_thershold = 0.05
cfg$lfc_threshold = 1
cfg$rate1_threshold = 0.5
```

@@

```{r paths2}
## to be modified when rendered
i = 2
filedir <- filelist[i] 
sample <- gsub(paste0(data_dir, "/"), "", filedir)

[3] We run DElegate::FindAllMarkers2 to find markers of the different clusters and manually check if they do make sense. 
DElegate::FindAllMarkers2 is an improved version of Seurat::FindAllMarkers based on pseudobulk differential expression method.

[4] We look at specific marker genes that we reported in the table marker.sets/CellType_metadata.csv to check the relevance of the clustering.

[5] We plot pca/umap reduction grouping with available annotations (singler_, cellassign_). 
We expect at least immune cells to be correctly label and fall into a few set of clusters.

[6] We run label transfer (Azimuth) to transfer annotation from the fetal kidney atlas human reference. 
We plot pca/umap reduction grouping with latest labels. We expect it to be the most representative of the cell types in the sample.

### Tested parameters

These other parameters have been tested but not selected:

- SCTransform (default parameters) --> RunPCA (default) --> RunUMAP (dims = 1:30) --> FindNeighbors (dims = 1:30) --> FindClusters (default)

## Input and output files

@@

@@

```{r seurat_wf, fig.width=8, fig.height=4, out.width='100%'}
# convert to seurat
srat <- CreateSeuratObject(counts = counts(sce),
                                    assay = "RNA",
                                    project = file)

@@

```{r fig.width=8, fig.height=4, out.width='100%'}

d2 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE) + ggtitle("Seurat Cluster - umap")
d1 <- SCpubr::do_DimPlot(srat, reduction="pca", group.by = "seurat_clusters", label = TRUE) + ggtitle("Seurat Cluster - pca")

d1|d2

@@

The umap clustering is expected for a triphasic Wilms tumor sample. The 3 main clusters might be:

- blastema cancer cells
- epithelium cancer and/or normal cells
- stroma cancer and/or normal cells

The 2 small clusters C11 and C13 might be immune and endothelial cells.

To be checked in the following analysis!

### Quick check of the QC values
```{r qc, fig.width=10, fig.height=2.5, out.width='100%'}
 SCpubr::do_ViolinPlot(srat, features = c("subsets_mito_percent"), ncol = 3, group.by = "seurat_clusters", legend.position = "none") 
```

@@

```{r fig.width=8, fig.height=4, out.width='100%'}

@@

d1|d2

@@

Cluster 3 might be composed of cycling cells, in higher proportion than the other clusters.

@@

```{r fig.width=20, fig.height=5, out.width='100%'}
 SCpubr::do_ViolinPlot(srat, features = rownames(srat)[rownames(srat) %in% (CellType_metadata$ENSEMBL_ID[CellType_metadata$cell_class %in% c("malignant")])] , ncol = 6, group.by = "seurat_clusters", legend.position = "none") 
```
The set of clusters 0-1-3-4-6-10-12-14 is NCAM1+ (ENSG00000149294) which is a marker of blastema area.
C2 is positive for COL6A3 which is a marker of stroma Wilms tumor cells (also normal stroma)

```{r fig.width=20, fig.height=15, out.width='100%'}
 SCpubr::do_FeaturePlot(srat, features = rownames(srat)[rownames(srat) %in% (CellType_metadata$ENSEMBL_ID[CellType_metadata$cell_class %in% c("malignant")])])
```

@@

```{r fig.width=20, fig.height=15, out.width='100%'}
 SCpubr::do_FeaturePlot(srat, features = rownames(srat)[rownames(srat) %in% (CellType_metadata$ENSEMBL_ID[CellType_metadata$cell_class %in% c("immune")])])
```
C11 is PTPRC+ (ENSG00000081237) and must be a cluster of immune cells. Wilms tumor are generally cold tumor, especially untreated Wilms tumor. According to the clinical data, the present sample is untreated (treatment = upfront resection). We do not expect high percentage of immune cells. The size of the cluster would therefore make sense.

@@

```{r fig.width=15, fig.height=5, out.width='100%'}
 SCpubr::do_FeaturePlot(srat, features = rownames(srat)[rownames(srat) %in% (CellType_metadata$ENSEMBL_ID[! CellType_metadata$cell_class %in% c("immune", "malignant")])], ncol = 3)
```

C13 is VWF+ (ENSG00000110799) and must be a cluster of endothelial cells.

@@

s.markers <- na.omit(s.markers)
s.markers %>%
    group_by(group1) %>%
    top_n(n =  5, wt = log_fc) -> top5

# subset for plotting
cells <- WhichCells(srat, downsample = 100)
ss <- subset(srat, cells = cells)
ss <- ScaleData(ss, c( top5$feature))

p1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE) + ggtitle("Seurat Cluster - umap")
p2 <- DoHeatmap(ss, features = top5$feature,  cells = cells, group.by = "seurat_clusters") + NoLegend() + 
  scale_fill_gradientn(colors =  c("#01665e","#35978f",'darkslategray3', "#f7f7f7", "#fee391","#fec44f","#F9AD03")) 
p3 <- ggplot(srat@meta.data, aes(seurat_clusters, fill = seurat_clusters)) + geom_bar() + NoLegend()
p4 <- VlnPlot(srat, features = c("nFeature_RNA", "nCount_RNA", "subsets_mito_percent"), ncol = 3)

common_title <- sprintf("Unsupervised clustering %s, %d cells", srat@meta.data$orig.ident[1], ncol(srat))
show((((p1 / p3) + plot_layout(heights = c(3,2)) | p2) / p4) + plot_layout(widths = c(1, 2)) + plot_layout(heights = c(3,1)) + plot_annotation(title = common_title))
DT::datatable(top5[,c(1,9,11,12)], caption = ("top 5 marker genes"), 
              extensions = 'Buttons', 
              options = list(  dom = 'Bfrtip',

@@

UNC5D (C1, dependency receptor), has been described in cancer and shown to regulate p53-dependent apoptosis (https://pubmed.ncbi.nlm.nih.gov/24691657/). 
We know that especially in blastema cell, p53 signaling play an important role.

PDCH7 (C7, protocadherin 7) is expressed in human kidney renal tubule (normal kidney epithelium, see https://www.proteinatlas.org/ENSG00000169851-PCDH7/tissue/kidney) and has been associated with oncogenic function in lung cancer (https://pubmed.ncbi.nlm.nih.gov/30409919/, https://www.frontiersin.org/journals/pharmacology/articles/10.3389/fphar.2023.1217213/full).

EBF1 (C8, Early B Cell factor 1) has been shown to play a role in kidney development, esp. glomerular, while expressed in stroma progenitor (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6727263/).

ROR1 (C8, Receptor tyrosine kinase-like orphan receptor 1) ROR1 is only detectable in embryonic tissue and generally absent in mature tissue. It has been associated with cancer (https://academic.oup.com/proteincell/article/5/7/496/6831810) and cancer stroma (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5367847/).

ERBB4 (C9, EGF receptor family) is expressed during kidney development of epithelial tubule and uteric bud (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3269924/)

FLT1 (C13, vascular endothelial growth factor receptor 1), PCAM and ELMO1 in the top5 of C13 marker genes confirm the identification of endothelial cells (https://www.ahajournals.org/doi/pdf/10.1161/CIRCRESAHA.109.213983, https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4986701/)

@@

Even if the result of SingleR and CellAssign are not specific to a Wilms tumor dataset, it confirm once again that C13 is a cluster of endothelial cells and C11 a cluster of immune cells.

@@

Note to the DataLab : where should I save the fetal kidney reference? 
In S3 like results as it is a quite big object or in the folder marker-sets?
The reference in now in the module folder, marker-sets
For more information related to the reference, please go to https://www.kidneycellatlas.org/

You will find:

@@

@@

# Preprocess with SCTransform
srat <-  RunAzimuth(srat, ref_dir, assay = 'RNA')
rm(reference)
gc()

DT::datatable(table(srat$seurat_clusters, srat$predicted.compartment), caption = ("table of azimuth compartment annotation per seurat clusters"),               extensions = 'Buttons', 
              options = list(  dom = 'Bfrtip',
              buttons = c( 'csv', 'excel')))

DT::datatable(table(srat$seurat_clusters, srat$predicted.celltype), caption = ("table of azimuth annotation per seurat clusters"), 
              extensions = 'Buttons', 
              options = list(  dom = 'Bfrtip',
              buttons = c( 'csv', 'excel')))

@@

d1 <- DimPlot(srat, reduction = "umap", group.by = "seurat_clusters", label = TRUE) + NoLegend()
d2 <- DimPlot(srat, reduction = "umap", dims = c(1,2), group.by = "predicted.compartment", label = TRUE, repel = TRUE) + NoLegend()
d3 <- DimPlot(srat, reduction = "umap", dims = c(1,2), group.by = "predicted.celltype", label = TRUE, repel = TRUE) + NoLegend()

@@

Looking at the predicted cell types, we observed that the set of cluster [0,1,3,4,6,10,12,14] is labeled as cap mesenchyme, an early developmental stage of nephron, while the set of clusters [5,7,9] is labeled with more differentiated epithelial units of the kidney. 
Based on this observation and the expression of NCAM1, UNC5D (C1) and PDCH7 (C7), ERBB4 (C9), we can conclude that the set of cluster [0,1,3,4,6,10,12,14] is mostly blastema cancer cells and the set of clusters [5,7,9] is eipthelial cancer or normal cells.

@@

f1 <- SCpubr::do_BarPlot(sample = srat,
                         group.by = "predicted.compartment",
                         split.by = "seurat_clusters",
                         position = "fill",
                         font.size = 10,
                         legend.ncol = 3) +
                         ggtitle("% cells")+
                         xlab(file)

f2 <- SCpubr::do_BarPlot(sample = srat,
                         group.by = "predicted.celltype",
                         split.by = "seurat_clusters",
                         position = "fill",
                         font.size = 10,
                         legend.ncol = 3) +
                         ggtitle("% cells")+
                         xlab(file)

@@

- C13 = endothelium (with high confidency)
- C11 = immune cells (with high confidency)

- C2 and C8, that form a unit in the umap reduction, are stromal cells, either cancer or normal (with high confidency)
- C[0-1-3-4-6-10-12-14] that form a unit in the umap reduction, are label as cap mesenchyme (<=> stem) and NCAM+ must be blastema cancer cells.
- C[5-7-9] that form a unit in the umap reduction, are label as epithelial cells (SingleR) and differentiated kidney epithelium subunit (Fetal kidney atlas) must be epithelial (cancer or normal)

The primary aim of this script was to assess the quality of clustering and play with clustering parameters. 
On this sample, we are happy with the clustering, as we could identify:
- one cluster of immune cells (cell marker PTPRC, SingleR, Azimuth label transfer fro the fetal kidney) 
- one cluster of endothelial cells (cell marker VWF, SingleR, Azimuth label transfer fro the fetal kidney)
- one set of clusters of stromal cells
- one set of cluster of epithelial cells
- one set of cluster of blastema cancer cells.

>>>>

---
title: "Evaluation of clustering parameters - SCPCS000168"
author: maud-p
date: '`r format(Sys.time(), "%B %d, %Y %H:%M:%S %Z")`'
output:
  html_document: 
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: pygments
    df_print: kable
params:
  sample_name: ~
  sample_path: ~
  out_rds_path: ~
---

@@

In this notebook, we will set up parameters in the seurat workflow [normalization --> reduction --> clustering] for one Wilms tumor sample (SCPCS000168) of the Wilms Tumor dataset (SCPCP000006).

@@

library(Azimuth)     # For label transfer

library(msigdbr)     # for gene set enrichment analysis or pathway enrichment
library(clusterProfiler)

library(data.table)
```

@@

```{r cfg}
cfg = list()
cfg$padj_thershold = 0.05
cfg$lfc_threshold = 1
cfg$rate1_threshold = 0.5
set.seed(12345)
```

@@

@@

```{r paths2}
## to be modified when rendered
i = 1
filedir <- filelist[i] 
sample <- gsub(paste0(data_dir, "/"), "", filedir)

Here we have a sample of anaplastic histology, untreated. Few thoughts:

- We thus expect anaplasia-associated patterns (TP53, MYCN as an example) to be major driver here.

- We do not expect high immune infiltration (induced by chemotherapy).

- We do not expect chemotherapy induced DNA damages, but maybe some DNA-repair pathways up in anaplstic cells.

[3] We look at specific marker genes that we reported in the table marker.sets/CellType_metadata.csv to check the relevance of the clustering.

[4] We look at specific pathways that we reported in the table marker.sets/Pathways_metadata.csv to check the relevance of the clustering.

[5] We run DElegate::FindAllMarkers2 to find markers of the different clusters and manually check if they do make sense. 
DElegate::FindAllMarkers2 is an improved version of Seurat::FindAllMarkers based on pseudobulk differential expression method.

[6] We perform enrichment analysis of marker genes for each seurat clusters. 
We defined all the genes from the seurat object as the universe and used the MSigDB gene sets.

[7] We plot pca/umap reduction grouping with available annotations (singler_, cellassign_). 
We expect at least immune cells to be correctly label and fall into a few set of clusters.

[8] We run label transfer (Azimuth) to transfer annotation from the fetal kidney atlas human reference. 
We plot pca/umap reduction grouping with latest labels. 
We expect it to be the most representative of the cell types in the sample.

Note: Other parameters have been previously tested, but we would like to show in the following report that the one selected is performing good.

@@

```{r seurat_wf, fig.width=8, fig.height=4, out.width='100%'}
# convert to seurat
srat <- CreateSeuratObject(counts = counts(sce),
                                    assay = "RNA",
                                    project = sample
                           )

@@

```{r fig.width=20, fig.height=5, out.width='100%'}

d2 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE) + ggtitle("Seurat Cluster - umap")
d1 <- SCpubr::do_DimPlot(srat, reduction="pca", group.by = "seurat_clusters", label = TRUE) + ggtitle("Seurat Cluster - pca")
v <- SCpubr::do_ViolinPlot(srat, features = c( "subsets_mito_percent"), ncol = 1, group.by = "seurat_clusters", legend.position = "none")

d1 + d2 + v + plot_layout(ncol = 3, widths = c(2,2,4))

@@

@@

```{r fig.width=20, fig.height=5, out.width='100%'}

@@

b <- SCpubr::do_BarPlot(sample = srat,
                         group.by = "Phase",
                         split.by = "seurat_clusters",
                         position = "fill",
                         font.size = 10,
                         legend.ncol = 3) +
                         ggtitle("% cells")+
                         xlab(file)
d1 + d2 + b + plot_layout(ncol = 3, widths = c(2,2,4))

@@

@@

```{r fig.width=25, fig.height=5, out.width='100%'}
 SCpubr::do_ViolinPlot(srat, features = rownames(srat)[rownames(srat) %in% (CellType_metadata$ENSEMBL_ID[CellType_metadata$cell_class %in% c("malignant")])] , ncol = 9, group.by = "seurat_clusters", legend.position = "none") 
```
All clustersexcept C13 are IGF2+ (ENSG00000167244).

```{r fig.width=20, fig.height=5, out.width='100%'}
d1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
d2 <- SCpubr::do_FeaturePlot(srat, features = "ENSG00000167244", pt.size = 0.2) + ggtitle("IGF2")
v <- SCpubr::do_ViolinPlot(srat, features = "ENSG00000167244", ncol = 1, group.by = "seurat_clusters", legend.position = "none")

d1 + d2 + v + plot_layout(ncol = 3, widths = c(1,1,3))
```
The set of clusters 0-1-4-8-9-11-12 is NCAM1+ (ENSG00000149294), which is a blastemal marker.

```{r fig.width=20, fig.height=5, out.width='100%'}
d1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
d2 <- SCpubr::do_FeaturePlot(srat, features = "ENSG00000149294",  pt.size = 0.2) + ggtitle("NCAM1")
v <- SCpubr::do_ViolinPlot(srat, features = "ENSG00000149294", ncol = 1, group.by = "seurat_clusters", legend.position = "none")

d1 + d2 + v + plot_layout(ncol = 3, widths = c(1,1,3))
```

All clusters except of C13 are WT1+ (ENSG00000184937).

```{r fig.width=20, fig.height=5, out.width='100%'}
d1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
d2 <- SCpubr::do_FeaturePlot(srat, features = "ENSG00000184937", pt.size = 0.2) + ggtitle("WT1")
v <- SCpubr::do_ViolinPlot(srat, features = "ENSG00000184937", ncol = 1, group.by = "seurat_clusters", legend.position = "none")

d1 + d2 + v + plot_layout(ncol = 3, widths = c(1,1,3))
```

C10 is positive for COL6A3 which is a marker of stroma Wilms tumor cells (also normal stroma)

```{r fig.width=20, fig.height=5, out.width='100%'}
d1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
d2 <- SCpubr::do_FeaturePlot(srat, features = "ENSG00000163359", pt.size = 0.2) + ggtitle("COL6A3")
v <- SCpubr::do_ViolinPlot(srat, features = "ENSG00000163359", ncol = 1, group.by = "seurat_clusters", legend.position = "none")

d1 + d2 + v + plot_layout(ncol = 3, widths = c(1,1,3))

@@

C13 is PTPRC+ (ENSG00000081237) and must be a cluster of immune cells. Wilms tumor are generally cold tumor, especially untreated Wilms tumor. According to the clinical data, the present sample is untreated (treatment = upfront resection). We do not expect high percentage of immune cells. The size of the cluster would therefore make sense.

```{r fig.width=20, fig.height=5, out.width='100%'}
d1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
d2 <- SCpubr::do_FeaturePlot(srat, features = "ENSG00000081237", pt.size = 0.2) + ggtitle("PTPRC = CD45")
v <- SCpubr::do_ViolinPlot(srat, features = "ENSG00000081237", ncol = 1, group.by = "seurat_clusters", legend.position = "none", pt.size = 2)

d1 + d2 + v + plot_layout(ncol = 3, widths = c(1,1,3))
```

@@

The set of clusters 8-11 is EPCAM+ (ENSG00000119888), which is a marker of epithelial (malignant and normal) kidney cells.

```{r fig.width=20, fig.height=5, out.width='100%'}
d1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
d2 <- SCpubr::do_FeaturePlot(srat, features = "ENSG00000119888", pt.size = 0.2) + ggtitle("EPCAM")
v <- SCpubr::do_ViolinPlot(srat, features = "ENSG00000119888", ncol = 1, group.by = "seurat_clusters", legend.position = "none")

d1 + d2 + v + plot_layout(ncol = 3, widths = c(1,1,3))
```

### Look at specific pathways

Here, we open the table of marker genes marker-sets/Pathway_metadata.csv.

```{r marker_pathways, fig.width=12, fig.height=6, out.width='100%'}
# open the set of reference marker genes

Pathway_metadata <- read.csv("marker-sets/Pathway_metadata.csv", header = TRUE, sep = ",")

DT::datatable(Pathway_metadata, caption = ("CellType_metadata"), 
              extensions = 'Buttons', 
              options = list(  dom = 'Bfrtip',

                               buttons = c( 'csv', 'excel')))
```
#### TP53 pathway

here we will calculate a TP53 score using AddMduleScore and the genes of the HALLMARK_P53_PATHWAY gene set.

```{r fig.width=10, fig.height=10, out.width='100%'}

## define genesets
hallmarks <- msigdbr(species = "human", category = "H")

TP53_list = hallmarks %>% 
  filter(gs_name == "HALLMARK_P53_PATHWAY") %>%
  dplyr::distinct(gs_name, gene_symbol, human_ensembl_gene) %>% as.data.frame()

TP53_list <- list(TP53_list$human_ensembl_gene)

srat <- AddModuleScore(srat, features = TP53_list, name = "TP53_score")
```

```{r fig.width=20, fig.height=5, out.width='100%'}
d1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
d2 <- SCpubr::do_FeaturePlot(srat, features = "TP53_score1", pt.size = 0.2) + ggtitle("TP53 score")
v <- SCpubr::do_ViolinPlot(srat, features = "TP53_score1", ncol = 1, group.by = "seurat_clusters", legend.position = "none")

d1 + d2 + v + plot_layout(ncol = 3, widths = c(1,1,3))
```
"Normal" cells (immune, C13) have a slightly higher TP53 score.

#### DNA repair pathway

here we will calculate a DNA_repair score using AddMduleScore and the genes of the HALLMARK_DNA_REPAIR gene set.

```{r fig.width=10, fig.height=10, out.width='100%'}

DNA_repair = hallmarks %>% 
  filter(gs_name == "HALLMARK_DNA_REPAIR") %>%
  dplyr::distinct(gs_name, gene_symbol, human_ensembl_gene) %>% as.data.frame()

DNA_repair_list <- list(DNA_repair$human_ensembl_gene)

srat <- AddModuleScore(srat, features = DNA_repair_list, name = "DNA_repair_score")
```

```{r fig.width=20, fig.height=5, out.width='100%'}
d1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
d2 <- SCpubr::do_FeaturePlot(srat, features = "DNA_repair_score1", pt.size = 0.2) + ggtitle("DNA_repair score")
v <- SCpubr::do_ViolinPlot(srat, features = "DNA_repair_score1", ncol = 1, group.by = "seurat_clusters", legend.position = "none")

d1 + d2 + v + plot_layout(ncol = 3, widths = c(1,1,3))
```
I am not sure how relevant the DNA repair score is. To be tested on more samples.

#### DROSHA target genes

```{r fig.width=10, fig.height=10, out.width='100%'}

## define genesets
c3 <- msigdbr(species = "human", category = "C3", subcategory = "TFT:GTRD")

DROSHA_list = c3 %>% 
  filter(gs_name == "DROSHA_TARGET_GENES") %>%
  dplyr::distinct(gs_name, gene_symbol, human_ensembl_gene) %>% as.data.frame()

DROSHA_list <- list(DROSHA_list$human_ensembl_gene)

srat <- AddModuleScore(srat, features = DROSHA_list, name = "DROSHA_score")
```

```{r fig.width=20, fig.height=5, out.width='100%'}
d1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
d2 <- SCpubr::do_FeaturePlot(srat, features = "DROSHA_score1", pt.size = 0.2) + ggtitle("DROSHA score")
v <- SCpubr::do_ViolinPlot(srat, features = "DROSHA_score1", ncol = 1, group.by = "seurat_clusters", legend.position = "none")

d1 + d2 + v + plot_layout(ncol = 3, widths = c(1,1,3))
```

#### DICER1 target genes

```{r fig.width=10, fig.height=10, out.width='100%'}
DICER1_list = c3 %>% 
  filter(gs_name == "DICER1_TARGET_GENES") %>%
  dplyr::distinct(gs_name, gene_symbol, human_ensembl_gene) %>% as.data.frame()

DICER1_list <- list(DICER1_list$human_ensembl_gene)

srat <- AddModuleScore(srat, features = DICER1_list, name = "DICER1_score")
```

```{r fig.width=20, fig.height=5, out.width='100%'}
d1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
d2 <- SCpubr::do_FeaturePlot(srat, features = "DICER1_score1", pt.size = 0.2) + ggtitle("DICER1 score")
v <- SCpubr::do_ViolinPlot(srat, features = "DICER1_score1", ncol = 1, group.by = "seurat_clusters", legend.position = "none")

d1 + d2 + v + plot_layout(ncol = 3, widths = c(1,1,3))
```
Not sure what to do with these scores.

@@

# Select top 5 genes for heatmap plotting
s.markers <- na.omit(s.markers)
s.markers %>%
    group_by(group1) %>%
    top_n(n =  5, wt = log_fc) -> top5

# subset for plotting
cells <- WhichCells(srat, downsample = 100)
ss <- subset(srat, cells = cells)
ss <- ScaleData(ss, features = top5$feature)

p1 <- SCpubr::do_DimPlot(srat, reduction="umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE) + ggtitle("Seurat Cluster - umap")
p2 <- DoHeatmap(ss, features = top5$feature,  cells = cells, group.by = "seurat_clusters") + NoLegend() + 
  scale_fill_gradientn(colors =  c("#01665e","#35978f",'darkslategray3', "#f7f7f7", "#fee391","#fec44f","#F9AD03")) 
p3 <- ggplot(srat@meta.data, aes(seurat_clusters, fill = seurat_clusters)) + geom_bar() + NoLegend()

common_title <- sprintf("Unsupervised clustering %s, %d cells", srat@meta.data$orig.ident[1], ncol(srat))
show((((p1 / p3) + plot_layout(heights = c(3,2)) | p2) ) + plot_layout(widths = c(1, 2)) + plot_layout(heights = c(3,1)) + plot_annotation(title = common_title))
DT::datatable(top5[, c(1, 9, 11, 12)], caption = ("top 5 marker genes"), 
              extensions = 'Buttons', 
              options = list(  dom = 'Bfrtip',

@@

PTPRC (C13, CD45), define the immune cluster.

### Enrichment analysis of marker genes

Here we perform enrichment analysis of the marker genes found in the previous section for each Seurat cluster.

We defined as universe/background all the genes expressed in the dataset, meaning the rownames of the Seurat object.

We used three gene sets from MSiGDB:

- the hallmark gene sets  are coherently expressed signatures derived by aggregating many MSigDB gene sets to represent well-defined biological states or processes.
- the C3 : regulatory target gene sets  based on gene target predictions for microRNA seed sequences and predicted transcription factor binding sites.
- the C8 : cell type signature gene sets  curated from cluster markers identified in single-cell sequencing studies of human tissue.

We used enricher function from clusterProfiler to perform enrichment analysis.

```{r fig.width=10, fig.height=10, out.width='100%'}
# define background genes = universe for enrichment
background <-  row_metadata$gene_ids

## define genesets
c8 <- msigdbr(species = "human", category = "C8")

msigdbr_hallmarks = hallmarks %>% dplyr::distinct(gs_name, ensembl_gene) %>% as.data.frame()
msigdbr_c3 = c3 %>% dplyr::distinct(gs_name, ensembl_gene) %>% as.data.frame()
msigdbr_c8 = c8 %>% dplyr::distinct(gs_name, ensembl_gene) %>% as.data.frame()
```

#### Hallmarks MSiGDB gene sets

```{r fig.width=25, fig.height=6, out.width='100%'}
tmp_H <- NULL
for(i in unique(srat$seurat_clusters)){
signature = s.markers$feature[s.markers$group1 == i]
ego_module <- enricher(gene = signature, universe = background, TERM2GENE = msigdbr_hallmarks)
tmp <- NULL

  if(!is.null(ego_module)){
    if(dim(ego_module)[1]>0){
    b_H <- barplot(ego_module, showCategory = 15, font.size = 20, label_format = 40)  
    tmp <- b_H$data
    tmp$set <- "Hallmarks"
    tmp$cluster <- i
    }
  }
tmp_H <- rbind(tmp_H, tmp)
}

ggplot(tmp_H, aes(Count, ID)) +
    geom_bar(stat = "identity", aes(fill=cluster))+

  theme_classic()+ 
    geom_text(
        aes(label = (paste( "Padj = ", round(p.adjust,2)))),
        color = "black",
        size = 3,
        hjust=1,
        position = position_dodge(0.5)
    )+
   theme(text = element_text(size=14))+
  facet_wrap(facets=c("cluster"), ncol = length(unique(tmp_H$cluster)))
```

MYC enrichment is expected in anaplastic samples.

#### C3 MSiGDB regulatory target gene sets

```{r fig.width=25, fig.height=6, out.width='100%'}

tmp_H <- NULL
for(i in unique(srat$seurat_clusters)){
signature = s.markers$feature[s.markers$group1 == i]
ego_module <- enricher(gene = signature, universe = background, TERM2GENE = msigdbr_c3)
tmp <- NULL

  if(!is.null(ego_module)){
    if(dim(ego_module)[1]>0){
    b_H <- barplot(ego_module, showCategory = 15, font.size = 20, label_format = 40)  
    tmp <- b_H$data
    tmp$ID <- factor(tmp$ID, levels = c(tmp$ID[order(tmp$Count)]))
    tmp$set <- "c8"
    tmp$cluster <- i
    }
  }
tmp_H <- rbind(tmp_H, tmp[1:5,])
}

ggplot(tmp_H, aes(Count, ID)) +
    geom_bar(stat = "identity", aes(fill=cluster))+

  theme_classic()+ 
    geom_text(
        aes(label = (paste( "Padj = ", round(p.adjust,2)))),
        color = "black",
        size = 3,
        hjust=1,
        position = position_dodge(0.5)
    )+
   theme(text = element_text(size=14))+
  facet_wrap(facets=c("cluster"), ncol = length(unique(tmp_H$cluster)))
```

#### C8 MSiGDB cell type signature gene sets

```{r fig.width=30, fig.height=15, out.width='100%'}

tmp_H <- NULL
for(i in unique(srat$seurat_clusters)){
signature = s.markers$feature[s.markers$group1 == i]
ego_module <- enricher(gene = signature, universe = background, TERM2GENE = msigdbr_c8)
tmp <- NULL

  if(!is.null(ego_module)){
    if(dim(ego_module)[1]>0){
    b_H <- barplot(ego_module, showCategory = 15, font.size = 20, label_format = 40)  
    tmp <- b_H$data
    tmp$ID <- factor(tmp$ID, levels = c(tmp$ID[order(tmp$Count)]))
    tmp$set <- "c8"
    tmp$cluster <- i
    }
  }
tmp_H <- rbind(tmp_H, tmp[1:5,])
}

ggplot(tmp_H, aes(Count, ID)) +
    geom_bar(stat = "identity", aes(fill=cluster))+

  theme_classic()+ 
    geom_text(
        aes(label = (paste( "Padj = ", round(p.adjust,2)))),
        color = "black",
        size = 3,
        hjust=1,
        position = position_dodge(0.5)
    )+
   theme(text = element_text(size=14))+
  facet_wrap(facets=c("cluster"), ncol = length(unique(tmp_H$cluster)))
```

- C10 is enriched in adulte differentiated cell types/pathway of the kidney.
This could suggest that C10 are from normal kidney. 
To be checked with inferCNV profile.

- C2, C5are enriched in muscle pathways, suggesting a stromal origin.

@@

Even if the result of SingleR and CellAssign are not specific to a Wilms tumor dataset, it give the first impression that C13 a cluster of immune cells.

@@

Note to the DataLab : where should I save the fetal kidney reference? 
In S3 like results as it is a quite big object or in the folder marker-sets?

For more information related to the reference, please go to https://www.kidneycellatlas.org/
You will find:

@@

This part is imho one of the most important step that allow us to have a quick and reliable idea of the composition of the different clusters. 
The predicted compartment are defined into 4 categories:

- endotheliúm
- stroma
- fetal nephron 
- immune

As for SingleR and CellAssign, the annotation of immune cells and endothelial cells is straightforward. 
The stroma compartment should then contain normal and cancer stromal cells.
The fetal nephron compartment contain blastema cancer cells a well as normal and cancer epithelial cells.

Further segregation of cancer versus normal cells will be achieved using a combination of markers/pathways (see above) and inferred CNV (to be done).

@@

# Preprocess with SCTransform
# Note, the RunAzimuth change a bit the Seurat object, changing for example the rownames to Symbol. 
# We thus choose to rename the object to keep the 
data <-  RunAzimuth(srat, ref_dir, assay = 'RNA')
rm(reference)
gc()

DT::datatable(table(data$seurat_clusters, data$predicted.compartment), caption = ("table of azimuth compartment annotation per seurat clusters"),               extensions = 'Buttons', 
              options = list(  dom = 'Bfrtip',
              buttons = c( 'csv', 'excel')))

DT::datatable(table(data$seurat_clusters, data$predicted.celltype), caption = ("table of azimuth annotation per seurat clusters"), 
              extensions = 'Buttons', 
              options = list(  dom = 'Bfrtip',
              buttons = c( 'csv', 'excel')))

@@

d1 <- DimPlot(data, reduction = "umap", group.by = "seurat_clusters", label = TRUE) + NoLegend()
d2 <- DimPlot(data, reduction = "umap", dims = c(1,2), group.by = "predicted.compartment", label = TRUE, repel = TRUE) + NoLegend()
d3 <- DimPlot(data, reduction = "umap", dims = c(1,2), group.by = "predicted.celltype", label = TRUE, repel = TRUE) + NoLegend()

@@

Looking at the predicted cell types, we observed that most of the clusters are labeled as cap mesenchyme, an early developmental stage of nephron.
This suggest that these clusters are either blastema or primitive epithelium.

@@

f1 <- SCpubr::do_BarPlot(sample = data,
                         group.by = "predicted.compartment",
                         split.by = "seurat_clusters",
                         position = "fill",
                         font.size = 10,
                         legend.ncol = 3) +
                         ggtitle("% cells")+
                         xlab(file)

f2 <- SCpubr::do_BarPlot(sample = data,
                         group.by = "predicted.celltype",
                         split.by = "seurat_clusters",
                         position = "fill",
                         font.size = 10,
                         legend.ncol = 3) +
                         ggtitle("% cells")+
                         xlab(file)

@@

- C13 = immune cells (with high confidency)

- C10, is composed of stroma (normal or cancer) cells and endothelium cells. Might be a cluster of normal cells, a bit mixed. To be checked with inferCNV.

- the two other set of clusters are cancer cells, with positivity for some Wilms tumor markers such as NCAM, WT1, EPCAM.
